---
title: "Lynx analysis"
author: "Benedikt Gehr, Stefanie Muff and Clara Panchaud"
date: "05/07/2021"
output: pdf_document
---


We start by loading the lynx dataset. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(survival)
library(glmmTMB)
library(TwoStepCLogit)
load("/Users/Clara/Documents/Master Thesis/Data/Lynx/data/lynx.RData")
dat <- lynx_table
```

First, we fit the models with fixed effects only. We want to check that the conditional logistic regression model gives the same results as the poisson reformulation.  We start by fitting the condition logistic regression model.
```{r}
run_clog=0
if (run_clog==1){
r.clogit <- clogit(formula=use~cover_swisstopo +altitude_swisstopo+
                     cover_swisstopo:ytcos2 +
                     cover_swisstopo:tsin +
                     cover_swisstopo:tsin2 +
                     cover_swisstopo:tcos2 +
                     altitude_swisstopo +
                     hum_indx +
                     prey_avail +
                     I(altitude_swisstopo^2) +
                     I(hum_indx^2) +
                     I(prey_avail^2) +
                     altitude_swisstopo:ytcos +
                   altitude_swisstopo:ytsin2 +
                   altitude_swisstopo:tcos +
                   altitude_swisstopo:tsin2 +
                   dist2 +
                   hum_indx_loc2:dist2 +
                   hum_indx:ytsin +
                   hum_indx:ytcos +
                   hum_indx:ytcos2 +
                   hum_indx:tsin +
                   hum_indx:tcos +
                   hum_indx:cover_swisstopo +
                   prey_avail:ytsin2 +
                   prey_avail:ytcos2 +
                   prey_avail:tsin +
                   strata(loc_id), data=dat) 

clog<-data.frame(summary(r.clogit)$coef)

write.csv(clog,"/Users/Clara/Documents/Master Thesis/Data/Lynx/clogit.csv",
          row.names = TRUE)
}
if (run_clog==0){
  clog=read.csv("/Users/Clara/Documents/Master Thesis/Data/Lynx/clogit.csv",
                row.names = 1)
}
```

Now, we fit the Poisson model and compare the results.

```{r,results=FALSE}
run_glmm=0
if (run_glmm==1){

TMBStruc = glmmTMB(use~cover_swisstopo +altitude_swisstopo+
                  cover_swisstopo:ytcos2 +
                   cover_swisstopo:tsin +
                   cover_swisstopo:tsin2 +
                   cover_swisstopo:tcos2 +
                   altitude_swisstopo +
                   hum_indx +
                   prey_avail +
                   I(altitude_swisstopo^2) +
                   I(hum_indx^2) +
                   I(prey_avail^2) +
                   altitude_swisstopo:ytcos +
                   altitude_swisstopo:ytsin2 +
                   altitude_swisstopo:tcos +
                   altitude_swisstopo:tsin2 +
                   dist2 +
                   hum_indx_loc2:dist2 +
                   hum_indx:ytsin +
                   hum_indx:ytcos +
                   hum_indx:ytcos2 +
                   hum_indx:tsin +
                   hum_indx:tcos +
                   hum_indx:cover_swisstopo +
                   prey_avail:ytsin2 +
                   prey_avail:ytcos2 +
                   prey_avail:tsin +
                   #prey_avail:tcos+
                   (1|loc_id), 
                   family=poisson,
                   data=dat,
                   doFit=FALSE)



TMBStruc$parameters$theta[1] = log(1e3)

TMBStruc$mapArg = list(theta=factor(c(NA)))
m1 = glmmTMB:::fitTMB(TMBStruc)
summary(m1)
glmm=summary(m1)$coef[1]
write.csv(glmm,"/Users/Clara/Documents/Master Thesis/Data/Lynx/glmm.csv",
          row.names = TRUE)
}
if (run_glmm==0){
  glmm=read.csv("/Users/Clara/Documents/Master Thesis/Data/Lynx/glmm.csv",
                row.names = 1)
}

round(clog[1],digit=2)==round(glmm[2:27,1],digit=2)
```

We can then move on to the random effects models. We start with the two-step estimation in order to fit a conditional logistic regression with random effects, which gives the same results as the paper that was published on this data.

```{r}
run_twostep=0

if (run_twostep==1){
lynx_model <-Ts.estim(formula = use~cover_swisstopo +
                        cover_swisstopo:ytcos2 +
                        cover_swisstopo:tsin +
                        cover_swisstopo:tsin2 +
                        cover_swisstopo:tcos2 +
                        altitude_swisstopo +
                        hum_indx +
                        prey_avail +
                        I(altitude_swisstopo^2) +
                        I(hum_indx^2) +
                        I(prey_avail^2) +
                        altitude_swisstopo:ytcos +
                        altitude_swisstopo:ytsin2 +
                        altitude_swisstopo:tcos +
                        altitude_swisstopo:tsin2 +
                        dist2 +
                        hum_indx_loc2:dist2 +
                        hum_indx:ytsin +
                        hum_indx:ytcos +
                        hum_indx:ytcos2 +
                        hum_indx:tsin +
                        hum_indx:tcos +
                        hum_indx:cover_swisstopo +
                        prey_avail:ytsin2 +
                        prey_avail:ytcos2 +
                        prey_avail:tsin +
                        prey_avail:tcos +
                        strata(loc_id)+cluster(id_anim), data = dat)
#lynx_model$r.effect
twostep<-cbind(beta=lynx_model$beta, se =lynx_model$se)

write.csv(twostep,"/Users/Clara/Documents/Master Thesis/Data/Lynx/twostep.csv",
          row.names = TRUE)
}
if (run_twostep==0){
  twostep=read.csv("/Users/Clara/Documents/Master Thesis/Data/Lynx/twostep.csv",
                   row.names = 1)
}
```

And finally, the Poisson model with random effects.

```{r}
run_random=0

if (run_random==1){
TMBStruc = glmmTMB(use~cover_swisstopo +altitude_swisstopo+
                     cover_swisstopo:ytcos2 +
                     cover_swisstopo:tsin +
                     cover_swisstopo:tsin2 +
                     cover_swisstopo:tcos2 +
                     hum_indx +
                     prey_avail +
                     I(altitude_swisstopo^2) +
                     I(hum_indx^2) +
                     I(prey_avail^2) +
                     altitude_swisstopo:ytcos +
                   altitude_swisstopo:ytsin2 +
                   altitude_swisstopo:tcos +
                   altitude_swisstopo:tsin2 +
                   dist2 +
                   hum_indx_loc2:dist2 +
                   hum_indx:ytsin +
                   hum_indx:ytcos +
                   hum_indx:ytcos2 +
                   hum_indx:tsin +
                   hum_indx:tcos +
                   hum_indx:cover_swisstopo +
                   prey_avail:ytsin2 +
                   prey_avail:ytcos2 +
                   prey_avail:tsin +
                   prey_avail:tcos+
                     (1|loc_id)+
                    (0+cover_swisstopo|id_anim) +
                     (0+altitude_swisstopo|id_anim)+
                     (0+hum_indx|id_anim)+
                     (0+prey_avail|id_anim)+
                     (0+cover_swisstopo:ytcos2|id_anim)+
                     (0+cover_swisstopo:tsin|id_anim)+
                     (0+cover_swisstopo:tsin2|id_anim)+
                     (0+cover_swisstopo:tcos2|id_anim)+
                     (0+I(altitude_swisstopo^2)|id_anim)+
                   (0+I(hum_indx^2)|id_anim)+
                   (0+I(prey_avail^2)|id_anim)+
                     (0+altitude_swisstopo:ytcos|id_anim)+
                   (0+altitude_swisstopo:ytsin2|id_anim)+
                   (0+altitude_swisstopo:tcos|id_anim)+
                   (0+altitude_swisstopo:tsin2|id_anim)+
                     (0+hum_indx_loc2:dist2|id_anim) +
                     (0+hum_indx:ytsin|id_anim) +
                     (0+hum_indx:ytcos|id_anim) +
                     (0+hum_indx:ytcos2|id_anim) +
                     (0+hum_indx:tsin|id_anim) +
                     (0+hum_indx:tcos|id_anim) +
                     (0+hum_indx:cover_swisstopo|id_anim) +
                     (0+prey_avail:ytsin2|id_anim) +
                     (0+prey_avail:ytcos2|id_anim) +
                     (0+prey_avail:tsin|id_anim) +
                     (0+prey_avail:tcos|id_anim) +
                   (0+dist2|id_anim), 
                   family=poisson,
                   data=dat,
                   doFit=FALSE)



TMBStruc$parameters$theta[1] = log(1e3)
TMBStruc$mapArg = list(theta=factor(c(NA,1:27)))
m = glmmTMB:::fitTMB(TMBStruc)

random=summary(m)$coefficients[1]
random
write.csv(random,"/Users/clara/Documents/Master Thesis/Data/Lynx/glmmrandomwithintercept.csv",
          row.names = TRUE)
}
if (run_random==0){
  random=read.csv("/Users/Clara/Documents/Master Thesis/Data/Lynx/glmmrandomwithintercept.csv",
                  row.names = 1)
}
```





